\scuba: iterative map-making with the Sub-Millimetre User Reduction Facility

### 2012-11-29 smurf_paper.tex
updates about lags in the bolo data, PCA section, and sky noise

\begin{abstract}
  The Submillimetre Common User Bolometer Array 2 (\scuba) is an
  instrument operating on the 15-m James Clerk Maxwell Telescope,
  nominally consisting of 5120 bolometers (with a useful yield of
  69\%) in each of two simultaneous imaging bands centred over 450 and
  850\,\micron. The camera is operated by scanning across the sky and
  recording data at a rate of 200\,Hz. As the largest of a new
  generation of multiplexed kilopixel bolometer cameras operating in
  the (sub)millimetre, \scuba\ data analysis represents a significant
  challenge.  We describe the production of maps using the
  Sub-Millimetre User Reduction Facility (SMURF) in which we have
  adopted a fast, iterative approach to map-making that enables data
  reduction on single, modern, high-end desktop computers, with
  execution times that are shorter than the observing times.  SMURF is
  used in an automated setting, both at the telescope for real-time
  feedback to observers, as well as for the production of science
  products for the JCMT Science Archive at the Canadian Astronomy Data
  Centre. Three detailed case studies are used to: (i) explore
  convergence properties of the map-maker using simple prior
  constraints (Uranus -- a point source); (ii) achieve the white-noise
  limit for faint point-source studies (extragalactic blank-field
  survey of the Lockman Hole); and (iii) demonstrate that our strategy
  is capable of recovering angular scales comparable to the size of
  the array footprint (approximately 5\,arcmin) for bright extended
  sources (star-forming region M17).
\end{abstract}

@@

In addition to uncorrelated white noise, and low-frequency (and often
correlated from bolometer to bolometer) $1/f$ drifts, bolometer power
spectra exhibit a roll-off at frequencies approaching Nyquist, which
is due to the anti-aliasing filter that forms part of the 200\,Hz
re-sampling stage. This anti-aliasing filter also introduces an
effective lag of $\sim7$\,ms (which must be added to the 1\,ms thermal
time constant of the bolometers mentioned above to obtain the total
lag). There are also line features which are thought to be produced at
frequencies far above the final 200\,Hz sample rate, and are aliased
to lower frequencies during the multiplexed readout stage before they
can be removed by the anti-aliasing filter.

@@

\caption{A comparison between single bolometer time-series in each of
  the 450 and 850\,\micron\ bands with the mixing chamber temperature,
  the line-of-sight 450\,\micron\ opacity derived from the water
  vapour monitor, and azimuth/elevation pointing offsets, before and
  after upgrading the instrument. The grey signals over-plotted in the
  top two panels show the residual time-series after removing the
  common-mode signals.  (a) Data taken during the S2SRO period,
  observation 29 on 20100313. There is a strong correlation between
  the bolometers and the roughly $\sim$25\,s oscillation in the
  fridge, but only a minor correlation with the opacity and telescope
  motion. The nearly flat common-mode subtracted signals show: (i)
  that most of the low-frequency signal is common to all of the
  bolometers; and (ii) the non-correlated, and predominantly white
  noise at 450\,\micron\ is significantly larger than at
  850\,\micron. (b) Data taken with the fully-commissioned instrument,
  observation 38 on 20111112. Unlike the S2SRO data, there is no
  strong signal produced by variations in the fridge temperature. The
  450\,\micron\ data have minimal correlated low-frequency noise (as
  evidenced by the lack of a difference between the raw and
  common-mode subtracted signals), although the 850\,\micron\ data
  show a common signal correlated with the opacity and telescope
  motion (note that this scan has a significantly larger amplitude
  than the S2SRO data set), mostly likely caused by a combination of
  changing airmass (anti-correlated with opacity), and magnetic field
  pickup (see Section~\ref{sec:magpickup} and
  Fig.~\ref{fig:magpickup}).}
\label{fig:bolos_mix}
\end{figure*}

@@

In Fig.~\ref{fig:bolos_mix} we show sample time-series from single
bolometers in each of the 450 and 850\,\micron\ focal planes, as well
as variations in the mixing chamber temperature (though not located in
the focal plane itself, it is certainly correlated with the
temperature of the detectors), and the telescope pointing, for two
data sets, before and after the upgrades that followed S2SRO.

In the S2SRO data (Fig.~\ref{fig:bolos_mix}a), observation 29 on
20100313, both bolometers share significant long-timescale structure
($\gsim10$\,s) that appears to be related to variations in the fridge
base temperature, although the similarity is clearly greater at
850\,\micron. In this particular case, the total power in the
fluctuations at 450\,\micron, are comparable to those at
850\,\micron. Such behaviour might be expected if there is a
comparable varying thermal load from the fridge at each wavelength
that dominates. We also note that there is no obvious strong
correlation between the low-frequency signal structure, at either
wavelength, with the telescope motion.

@@

Data from the fully commissioned instrument, observation 38 on
20111112, are shown in Fig.~\ref{fig:bolos_mix}b. Having solved the
fridge oscillation problem, these data no longer exhibit a correlation
with the mixing chamber (however, note that variations are still seen
in the mixing chamber signal; such variations do not necessarily
reflect changes in the focal plane temperature). There is minimal
correlated low-frequency noise in these 450\,\micron\ data (note that
there is almost no difference once the common-mode has been
subtracted). The 850\,\micron\ channel, however, exhibits a more
significant signal that is obviously correlated with the telescope
motion. It is also correlated amongst many of the detectors, and
common-mode removal corrects it to a large extent. Part of the reason
that this is seen here, and not in the S2SRO data set shown, is that
the amplitude of the scan pattern is larger. Most of this
``scan-synchronous'' noise is attributed to magnetic field pickup, as
described in Section~\ref{sec:magpickup}.

@@

The basic method is as follows: (i) a covariance matrix is built up
for all pairs $(i,j)$ of the $N$ bolometer time-series,
$\langle\mathbf{b}_i(t),\mathbf{b}_j(t)\rangle$; and (ii) a singular
value decomposition identifies a new set of statistically uncorrelated
eigenvectors, $\mathbf{\xi}_i(t)$ (i.e., whose covariance matrix is
diagonal), such that each of the bolometer time-series is a linear
combination of the eigenvectors, or components i.e., $\mathbf{b}_i(t)
= \bar{\mathbf{\xi}} \mathbf{\lambda}_i^\mathrm{T}$, where each row of
the matrix $\bar{\mathbf{\xi}}$ is an eigenvector, and
$\mathbf{\lambda}_i^\mathrm{T}$ is a column vector containing the
corresponding eigenvalues. The $\mathbf{\xi}_i(t)$ are normalised by
$[\sum_t \mathbf{\xi}_i^2(t)]^{1/2}$, such that the root mean square
(\rms) amplitude of each component in a given bolometer signal is
stored in the eigenvalues. In the earlier analyses mentioned, the
low-frequency noise is assumed to be encapsulated in those components
with the largest eigenvalues. Removing the projection of the time
series along those components then significantly reduces $1/f$ noise
while retaining most of the (higher-frequency) signal in point-like
sources

A novel feature of our \scuba\ analysis is that we can perform PCA
with 450 and 850\,\micron\ data simultaneously, potentially helping us
to differentiate thermal and optical noise signals (e.g., from the
atmosphere) that might appear in both wavelengths, from other noise
sources that are restricted to single subarrays (such as readout
noise). In Fig.~\ref{fig:pca} we show the results for a combined
analysis of the s4a and s8b subarrays for the first $\sim$100\,s of
the same observation that was used in Figs.~\ref{fig:bolos_mix}b and
\ref{fig:pspec}b. The 12 most significant components are shown (ranked
by the mean eigenvalues across all working bolometers in both
bands). The top and middle panels for each component show the
time-series and PSDs of the normalised eigenvectors. The bottom panels
are maps indicating the eigenvalues (amplitudes) of the component
across the focal plane (with the mean $\bar{\lambda}$ and and \rms,
$\lambda_\mathrm{rms}$, of the eigenvalues for all of the bolometers
calculated separately at each wavelength also shown).

The majority of the correlated signal at both wavelengths in
Fig.~\ref{fig:pca} is clearly produced by Components~1 and 2. In both
cases, the eigenvector time-series exhibit a roughly periodic signal
that resembles the scan pattern in Fig.~\ref{fig:bolos_mix}b. While
these correlated signals also clearly appear at both wavelengths,
referring to the maps of eigenvalues, they are considerably stronger
at 850\,\micron\ (consistent with the visual appearance of the
bolometer signals in Fig.~\ref{fig:pca}). Also note that the PSDs for
these eigenvectors exhibit nearly a pure $1/f$ signature with almost
no line features, or a white-noise plateau (suggesting that these are
high-\snr\ measurements of a purely low-frequency drift).

@@

Why are there two dominant components? First, the smoothly-varying
pattern in the eigenvalue maps (particularly evident at 850\,\micron)
suggests that the scan-synchronous noise (presumably magnetic field
pickup) has a different response across the focal plane; in this case
the PCA may have identified two orthogonal shapes that, when mixed in
different quantities, can reproduce most of the signal in each
bolometer. It is also likely that a varying atmospheric contribution
is also included in these eigenvectors. One would expect the
atmosphere to be more significant at 450 than at 850\,\micron,
although it is not obviously visible due to the strong presence of the
scan-synchronous noise.

Next, Components~3, 4, 7 and 11 in Fig.~\ref{fig:pca} have a common
character. In all cases, the PCA appears to have identified strong,
but transient glitches that appear in only one or several bolometers,
such as the two black bolometers near column 8, row 21 at
850\,\micron\ in Component~3. Such events are typically flagged and
removed during iterative de-spiking (Section~\ref{sec:ast}).

The most interesting feature of Component~5 in Fig.~\ref{fig:pca} is a
strong narrow line in the PSD at a frequency of about 57\,Hz
(suggesting the 60\,Hz mains as the culprit). The eigenvalue maps show
that it is only significant for several bolometers in the
850\,\micron\ subarray: at column 22 row 22; several bolometers at the
bottom of column 28; and more near the top-right of the subarray.

Components~6 and 8--10 in Fig.~\ref{fig:pca} have several generic
features in common. First, the PSDs exhibit moderate low-frequency
$1/f$ noise. Second, there are broad-line features at frequencies
above $\sim$30\,Hz which we believe to be aliased high-frequency
noise, in part due to the clear correlation patterns that appear along
columns in the eigenvalue maps at both wavelengths (e.g., columns 26
and 29 at 450\,\micron\ in Component~6). Finally, these components
have potentially unrelated features mixed into them, like sudden
offsets that are likely residuals due to imperfect step correction
(e.g., around $t\sim50$\,s in the eigenvector time-series for
Component~6; also see Section~\ref{sec:steps}).

Finally, Component~12 in Fig.~\ref{fig:pca} is significant in only a
handful of detectors in the 850\,\micron\ subarray (e.g., column 21
row 25). The time-series plot for the eigenvector exhibits a large
positive tail of sudden excursions, which gradually worsen over
time. Bolometers that exhibit this signal are probably poorly biased,
and are usually flagged due to elevated noise levels during
pre-processing (Section~\ref{sec:flagbadbol}).

This example illustrates some of the types of correlated signals and
patterns that PCA can identify in \scuba\ data. While there are
typically one or two strong signal components detected, in general,
the details can vary significantly from data set to data set, and the
lengths of the time-series analysed. The reason for this is probably
due to the fact that many of the noise sources are not stationary in
time (e.g., scan-synchronous noise which obviously depends on the scan
pattern, and the tuning of the subarrays). It should also be clear
from this example that while PCA offers some helpful insight into the
various sources of noise, it does not necessarily identify clean
patterns that can easily be modelled. For example, while there are
clearly correlation patterns along columns (Components~6 and 8--10,
particularly at 450\,\micron), as one might expect given the common
amplification chain, the intensities of these high-frequency signals
appear random in the eigenvalue maps, and a common-mode signal
estimated from the data for each column would not remove it. Finally,
we note that most of the high-frequency ($\gsim2$\,Hz) correlated
features that have been identified tend to be restricted to a small
number of detectors, and can often be identified and removed using
other (faster) means. What remains is predominantly low-frequency
noise. This conclusion is central to our data reduction strategy
described in the next section.

@@

The extinction correction is a multiplicative factor that is normally
derived using the JCMT water vapour monitor (WVM), and is not
considered to be a free parameter in the solution [$\mathbf{e}_i(t)$
in Eq.~\ref{eq:model}]. However, it is applied as part of the
iterative solution, rather than a pre-processing step, since any small
errors will be amplified by the large low-frequency drifts in the raw
bolometer time-series. For example, if the bolometer drift is 1000
times greater than an astronomical source of interest, a 1 per cent
error in the flatfield will produce stripes of order 10 times the
astronomical signal amplitude in the final map! If \model{EXT} is
applied after the bulk of the low-frequency noise has been removed
(e.g., \model{COM,GAI}), then there is little potential for such small
errors to affect the final map.  For details on how it is calculated
see \citet{dempsey2012}. Note that its numerical value is calculated
only once, and simply applied as a scale factor in the iterative
solution. Unlike the additive model components, it is inverted at the
start of each iteration to preserve the amplitude of the data before
the re-calculation of other model components.

>>>>

\begin{abstract}
  The Submillimetre Common User Bolometer Array 2 (\scuba) is an
  instrument operating on the 15-m James Clerk Maxwell Telescope,
  nominally consisting of 5120 bolometers (with a useful yield of
  69\%) in each of two simultaneous imaging bands centred over 450 and
  850\,\micron. The camera is operated by scanning across the sky and
  recording data at a rate of 200\,Hz. As the largest of a new
  generation of multiplexed kilopixel bolometer cameras operating in
  the (sub)millimetre, \scuba\ data analysis represents a significant
  challenge.  We describe the production of maps using the
  Sub-Millimetre User Reduction Facility (SMURF) in which we have
  adopted a fast, iterative approach to map-making that enables data
  reduction on single, modern, high-end desktop computers, with
  execution times that are typically shorter than the observing times.
  SMURF is used in an automated setting, both at the telescope for
  real-time feedback to observers, as well as for the production of
  science products for the JCMT Science Archive at the Canadian
  Astronomy Data Centre. Three detailed case studies are used to: (i)
  explore convergence properties of the map-maker using simple prior
  constraints (Uranus -- a point source); (ii) achieve the white-noise
  limit for faint point-source studies (extragalactic blank-field
  survey of the Lockman Hole); and (iii) demonstrate that our strategy
  is capable of recovering angular scales comparable to the size of
  the array footprint (approximately 5\,arcmin) for bright extended
  sources (star-forming region M17).
\end{abstract}

@@

In addition to uncorrelated white noise, and low-frequency (and often
correlated from bolometer to bolometer) $1/f$ drifts, bolometer power
spectra exhibit a roll-off at frequencies approaching Nyquist, which
is due to the anti-aliasing filter that forms part of the 200\,Hz
re-sampling stage. This anti-aliasing filter also introduces an
effective lag of $\sim$6\,ms. Adding this lag to the $\sim$1\,ms
thermal time constant, as well as a negative shift of about $-$2.5\,ms
caused by a half-sample offset due to the synchronization between the
pointing and bolometer data, yields a net lag of approximately
4.5\,ms. There are also line features in the spectra that are thought
to be produced primarily at frequencies far above the final 200\,Hz
sample rate, and are aliased to lower frequencies during the
multiplexed readout stage before they can be removed by the
anti-aliasing filter.

@@

\caption{A comparison between single bolometer time-series in each of
  the 450 and 850\,\micron\ bands with the mixing chamber temperature,
  the line-of-sight 450\,\micron\ opacity derived from the JCMT water
  vapour monitor, and azimuth/elevation pointing offsets, before and
  after upgrading the instrument. The grey signals over-plotted in the
  top two panels show the residual time-series after removing the
  common-mode signals.  (a) Data taken during the S2SRO period,
  observation 29 on 20100313. There is a strong correlation between
  the bolometers and the roughly $\sim$25\,s oscillation in the
  fridge, but only a minor correlation with the opacity and telescope
  motion. The nearly flat common-mode subtracted signals show: (i)
  that most of the low-frequency signal is common to all of the
  bolometers; and (ii) the non-correlated, and predominantly white
  noise at 450\,\micron\ is significantly larger than at
  850\,\micron. (b) Data taken with the fully-commissioned instrument,
  observation 38 on 20111112. Unlike the S2SRO data, there is no
  strong signal produced by variations in the fridge temperature. The
  450\,\micron\ data have minimal correlated low-frequency noise (as
  evidenced by the lack of a difference between the raw and
  common-mode subtracted signals), although the 850\,\micron\ data
  show a common signal correlated with the opacity and telescope
  motion (note that this scan has a significantly larger amplitude
  than the S2SRO data set), mostly likely caused by a combination of
  changing airmass (anti-correlated with opacity), and magnetic field
  pickup (see Section~\ref{sec:magpickup} and
  Fig.~\ref{fig:magpickup}).}
\label{fig:bolos_mix}
\end{figure*}

@@

In Fig.~\ref{fig:bolos_mix} we show sample time-series from single
bolometers in each of the 450 and 850\,\micron\ focal planes, as well
as variations in the mixing chamber temperature (though not located in
the focal plane itself, it is certainly correlated with the
temperature of the detectors), the line-of-sight atmoshperic opacity
as derived from the JCMT water vapour monitor (WVM), and the telescope
pointing, for two data sets, before and after the upgrades that
followed S2SRO.

In the S2SRO data (Fig.~\ref{fig:bolos_mix}a), observation 29 on
20100313, both bolometers share significant long-timescale structure
($\gsim10$\,s) that appears to be related to variations in the fridge
base temperature, although the similarity is clearly greater at
850\,\micron. Note that the total power in the fluctuations at
450\,\micron\ are comparable to those at 850\,\micron\ as one might
expect if there is a comparable varying thermal load from the fridge
at each wavelength that dominates. We also note that there is no
obvious strong correlation between the low-frequency signal structure,
at either wavelength, with the telescope motion. However, there is a
suggestion that the shorter-timescale behaviour of the opacity is
anti-correlated with the elevation, as expected.

@@

Data from the fully commissioned instrument, observation 38 on
20111112, are shown in Fig.~\ref{fig:bolos_mix}b. Having solved the
fridge oscillation problem, these data no longer exhibit a correlation
with the mixing chamber (however, note that variations are still seen
in the mixing chamber signal; such variations do not necessarily
reflect changes in the focal plane temperature). There is minimal
correlated low-frequency noise in these 450\,\micron\ data, resulting
in little difference between the raw and common-mode subtracted
data. The 850\,\micron\ channel, however, exhibits a more significant
signal that is obviously correlated with the telescope motion. It is
also correlated amongst many of the detectors, and common-mode removal
corrects it to a large extent. Part of the reason that this is seen
here, and not in the S2SRO data set shown, is that the amplitude of
the scan pattern is larger. Most of this ``scan-synchronous'' noise is
attributed to magnetic field pickup, as described in
Section~\ref{sec:magpickup}. Also similar to the S2SRO data, an
anti-correlation between the opacity and elevation is apparent.

@@

The basic method is as follows: (i) a covariance matrix is built up
for all pairs $(i,j)$ of the $N$ bolometer time-series,
$\langle\mathbf{b}_i(t),\mathbf{b}_j(t)\rangle$; and (ii) a singular
value decomposition identifies a new set of statistically uncorrelated
eigenvectors, $\mathbf{\xi}_i(t)$ (i.e., whose covariance matrix is
diagonal), such that each of the bolometer time-series is a linear
combination of the eigenvectors, or components i.e., $\mathbf{b}_i(t)
= \bar{\mathbf{\xi}} \mathbf{\lambda}_i^\mathrm{T}$, where each row of
the matrix $\bar{\mathbf{\xi}}$ is an eigenvector, and
$\mathbf{\lambda}_i^\mathrm{T}$ is a column vector containing the
corresponding eigenvalues. The $\mathbf{\xi}_i(t)$ are normalised by
$[\sum_t \mathbf{\xi}_i^2(t)]^{1/2}$, such that the amplitude of each
component in a given bolometer signal is stored in the eigenvalues. In
the earlier analyses mentioned, the low-frequency noise is assumed to
be encapsulated in those components with the largest
eigenvalues. Removing the projection of the time series along those
components then significantly reduces $1/f$ noise while retaining most
of the (higher-frequency) signal in point-like sources

A novel feature of our \scuba\ analysis is that we can perform PCA
with 450 and 850\,\micron\ data simultaneously, potentially helping us
to differentiate thermal and optical noise signals (e.g., from the
atmosphere) that might appear in both wavelengths, from other noise
sources that are restricted to single subarrays (such as readout
noise). In Fig.~\ref{fig:pca} we show the results for a combined
analysis of the s4a and s8b subarrays for the first $\sim$100\,s of
the same observation that was used in Figs.~\ref{fig:bolos_mix}b and
\ref{fig:pspec}b. The 9 most significant components are shown (ranked
by the mean eigenvalues across all working bolometers in both
bands). The top and middle panels for each component show the
time-series and PSDs of the normalised eigenvectors. The bottom panels
are maps indicating the eigenvalues (amplitudes) of the component
across the focal plane (with the mean $\bar{\lambda}$ and and \rms,
$\lambda_\mathrm{rms}$, of the eigenvalues for all of the bolometers
calculated separately at each wavelength also shown).

The majority of the correlated signal at both wavelengths in
Fig.~\ref{fig:pca} is produced by Component~1, in which the
eigenvector time-series exhibits a roughly periodic signal that
resembles the scan pattern in Fig.~\ref{fig:bolos_mix}b. While this
correlated signal appears at both wavelengths, referring to the maps
of eigenvalues, they are considerably stronger at 850\,\micron\
(consistent with the visual appearance of the bolometer signals in
Fig.~\ref{fig:pca}). Also note that the PSD for theis eigenvector
exhibit nearly a pure $1/f$ signature with minimal line features, or a
white-noise plateau (suggesting that these are high-\snr\ measurements
of a purely low-frequency drift). Component~3 is weaker, yet also
shows evidence of the scan-synchronous noise, combined with
high-frequency line features that appear primarily in columns 26, 29
and 30 of the 450\,\micron\ subarray. Unlike Component~1, it is
generally more dominant at 450\,\micron.

@@

Why are there two components that seem to be dominated by
scan-synchronous noise (which is presumably a mixture of
magnetic-field pickup and atmospheric noise)? First, the
smoothly-varying gradients in the eigenvalue maps (primarily at
850\,\micron) suggests that this noise has a different response across
the focal plane; in this case the PCA has identified two orthogonal
shapes that, when mixed in different quantities, can reproduce most of
the signal in each bolometer. It is also likely that the atmospheric
contribution is stronger at 450\,\micron\ (i.e., predominantly
Component~3).

Like Componen~1, Component~2 is a purely $1/f$ signal, with an
amplitude comparable to Component~3, that appears almost exclusively
at 850\,\micron. Since it is not obviously correlated with the
telescope motion, yet has an amplitude that varies smoothly across the
850\,\micron\ subarray, it may represent the PCAs attempt to reproduce
some non-linear response to magnetic-field and/or atmospheric pickup

The remaining, weaker, Components~4--9 share some common
features. They all have some degree of $1/f$ noise, suggesting further
residual noise related to the telescope motion and/or other slow
baseline drifts. All of them also exhibit line features in their
spectra, which in many cases appear at random amplitudes along columns
(e.g. columns 15, and 26--30 at 450\,\micron\ in Component~4), or in a
sampling of isolated bolometers (e.g. colummn 13, row 19 at
850\,\micron\ in Component~6). The signals that appear along columns
is probably related to magnetic-field pickup. Other sources of lines
are thought to be aliased high-frequency noise, and also the 60\,Hz
mains which appear in some bolometers.

This example illustrates some of the types of correlated signals and
patterns that PCA can identify in \scuba\ data. While there are
typically one or two strong signal components detected, in general,
the details can vary significantly from data set to data set, and the
lengths of the time-series analysed. The reason for this is that many
of the noise sources are not stationary in time (e.g.,
scan-synchronous noise which obviously depends on the scan pattern,
and the tuning of the subarrays). It should also be clear from this
example that while PCA offers some helpful insight into the various
sources of noise, it does not necessarily identify clean patterns that
can easily be modelled. For example, while there are clearly
correlation patterns along columns as expected from the common
amplification chain, the intensities of these high-frequency signals
appear random in the eigenvalue maps, and a simple common-mode signal
estimated from the data for each column would not remove
it. Regardless, the dominant noise, in all cases, is at
low-frequencies. This conclusion is central to our data reduction
strategy described in Section~\ref{sec:algorithm}.

\subsection{Overall performance and mapping speed}

@@

The extinction correction is a multiplicative factor that is normally
derived using the WVM, and is not considered to be a free parameter in
the solution [$\mathbf{e}_i(t)$ in Eq.~\ref{eq:model}]. However, it is
applied as part of the iterative solution, rather than a
pre-processing step, since any small errors will be amplified by the
large low-frequency drifts in the raw bolometer time-series. For
example, if the bolometer drift is 1000 times greater than an
astronomical source of interest, a 1 per cent error in the flatfield
will produce stripes of order 10 times the astronomical signal
amplitude in the final map! If \model{EXT} is applied after the bulk
of the low-frequency noise has been removed (e.g., \model{COM,GAI}),
then there is little potential for such small errors to affect the
final map.  For details on how it is calculated see
\citet{dempsey2012}. Note that its numerical value is calculated only
once, and simply applied as a scale factor in the iterative
solution. Unlike the additive model components, it is inverted at the
start of each iteration to preserve the amplitude of the data before
the re-calculation of other model components.
